 extracting environmenal and oceanographic variables

### 2023-10-16 03_environ_var.Rmd
Update 03_environ_var.Rmd

Required Packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(sf)
library(terra)
library(PAMpal)
library(PAMmisc)
library(rerddap)
```

Read in GPSwhale rdas
```{r}
env07 <- readRDS(file = "data/GPSwhaleAIS7.rda")
env08 <- readRDS(file = "data/GPSwhaleAIS8.rda")
env10 <- readRDS(file = "data/GPSwhaleAIS10.rda")
env12 <- readRDS(file = "data/GPSwhaleAIS12.rda")
env13 <- readRDS(file = "data/GPSwhaleAIS13.rda")
env14 <- readRDS(file = "data/GPSwhaleAIS14.rda")
env16 <- readRDS(file = "data/GPSwhaleAIS16.rda")
env18 <- readRDS(file = "data/GPSwhaleAIS18.rda")
env19 <- readRDS(file = "data/GPSwhaleAIS19.rda")
env20 <- readRDS(file = "data/GPSwhaleAIS20.rda")
env21 <- readRDS(file = "data/GPSwhaleAIS21.rda")
env22 <- readRDS(file = "data/GPSwhaleAIS22.rda")
env23 <- readRDS(file = "data/GPSwhaleAIS23.rda")
```

@@

Sea Surface Temperature from ERDDAP using PAMPal function
Daily Multi-scale Ultra-high Resolution (MUR) SST Analysis fv04.1, Global, 0.01°
```{r}
# Downloads data from servers onto external hard drive. Uncomment and only run once.
#env07 <- matchEnvData(GPSwhaleAIS7, nc = 'jplMURSST41', fileName = "D:/Environ_variables/analysedSST07", var = c('analysed_sst'))
#env08 <- matchEnvData(GPSwhaleAIS8, nc = 'jplMURSST41', fileName = "D:/Environ_variables/analysedSST08", var = c('analysed_sst'))
#env10 <- matchEnvData(GPSwhaleAIS10, nc = 'jplMURSST41', fileName = "D:/Environ_variables/analysedSST10", var = c('analysed_sst'))
#env12 <- matchEnvData(GPSwhaleAIS12, nc = 'jplMURSST41', fileName = "D:/Environ_variables/analysedSST12", var = c('analysed_sst'))
#env13 <- matchEnvData(GPSwhaleAIS13, nc = 'jplMURSST41', fileName = "D:/Environ_variables/analysedSST13", var = c('analysed_sst'))
#env14 <- matchEnvData(GPSwhaleAIS14, nc = 'jplMURSST41', fileName = "D:/Environ_variables/analysedSST14", var = c('analysed_sst'))
#env16 <- matchEnvData(GPSwhaleAIS16, nc = 'jplMURSST41', fileName = "D:/Environ_variables/analysedSST16", var = c('analysed_sst'))
#env18 <- matchEnvData(GPSwhaleAIS18, nc = 'jplMURSST41', fileName = "D:/Environ_variables/analysedSST18", var = c('analysed_sst'))
#env19 <- matchEnvData(GPSwhaleAIS19, nc = 'jplMURSST41', fileName = "D:/Environ_variables/analysedSST19", var = c('analysed_sst'))
#env20 <- matchEnvData(GPSwhaleAIS20, nc = 'jplMURSST41', fileName = "D:/Environ_variables/analysedSST20", var = c('analysed_sst'))
#env21 <- matchEnvData(GPSwhaleAIS21, nc = 'jplMURSST41', fileName = "D:/Environ_variables/analysedSST21", var = c('analysed_sst'))
#env22 <- matchEnvData(GPSwhaleAIS22, nc = 'jplMURSST41', fileName = "D:/Environ_variables/analysedSST22", var = c('analysed_sst'))
#env23 <- matchEnvData(GPSwhaleAIS23, nc = 'jplMURSST41', fileName = "D:/Environ_variables/analysedSST23", var = c('analysed_sst'))
```

@@

  assign(paste0("env", substr(ss, 33, 34)), data)
}
```
### Wind Stress Curl from ERDDAP using PAMPal function
Wind Stress, Metop-A ASCAT, 0.25°, Global, Near Real Time, 2009-present (1 Day) 
```{r}
# Downloads data from servers onto external hard drive. Uncomment and only run once.
#env07 <- matchEnvData(env07, nc = 'erdQAstress1day', fileName = "D:/Environ_variables/windstress07.nc", var = c('curl'))
#env08 <- matchEnvData(env08, nc = 'erdQAstress1day', fileName = "D:/Environ_variables/windstress08.nc", var = c('curl'))
#env10 <- matchEnvData(env10, nc = 'erdQAstress1day', fileName = "D:/Environ_variables/windstress10.nc", var = c('curl'))
#env12 <- matchEnvData(env12, nc = 'erdQAstress1day', fileName = "D:/Environ_variables/windstress12.nc", var = c('curl'))
#env13 <- matchEnvData(env13, nc = 'erdQAstress1day', fileName = "D:/Environ_variables/windstress13.nc", var = c('curl'))
#env14 <- matchEnvData(env14, nc = 'erdQAstress1day', fileName = "D:/Environ_variables/windstress14.nc", var = c('curl'))
#env16 <- matchEnvData(env16, nc = 'erdQAstress1day', fileName = "D:/Environ_variables/windstress16.nc", var = c('curl'))
#env18 <- matchEnvData(env18, nc = 'erdQAstress1day', fileName = "D:/Environ_variables/windstress18.nc", var = c('curl'))
#env19 <- matchEnvData(env19, nc = 'erdQAstress1day', fileName = "D:/Environ_variables/windstress19.nc", var = c('curl'))
#env20 <- matchEnvData(env20, nc = 'erdQAstress1day', fileName = "D:/Environ_variables/windstress20.nc", var = c('curl'))
#env21 <- matchEnvData(env21, nc = 'erdQAstress1day', fileName = "D:/Environ_variables/windstress21.nc", var = c('curl'))
#env22 <- matchEnvData(env22, nc = 'erdQAstress1day', fileName = "D:/Environ_variables/windstress22.nc", var = c('curl'))
#env23 <- matchEnvData(env23, nc = 'erdQAstress1day', fileName = "D:/Environ_variables/windstress23.nc", var = c('curl'))
```
Once downloaded, matches wind stress curl from file
```{r}
windlist <- dir("D:/Environ_variables", recursive=TRUE, full.names=TRUE, pattern="windstress")

for (w in windlist) {
  data <- matchEnvData(get(paste0("env", substr(w, 32, 33))), nc = w) %>%
     dplyr::select(-matchLong_mean, -matchLat_mean, -matchTime_mean, -matchDepth_mean)   # remove unnecessary columns

@@

### Distance to continental slope (uses terra and sf package)
```{r}
bathy <- rast("D:/CCES_gebco_bathy.tif") # Bathymetry for all of CCE from https://download.gebco.net/
agg <- terra::aggregate(bathy, 5)  # Increase cell size by factor of 5 to be visible

@@

tracklist <- grep(pattern = "env", names(.GlobalEnv), value = T)

@@

  assign(paste0("env", substr(t, 4, 5)), dbind)
}
```

Depth from ERDDAP using PAMPal function
SRTM15_PLUS Estimated Topography, 15 seconds, Global, v1 
```{r}
# Downloads data from servers onto external hard drive. Uncomment and only run once.
#env07 <- matchEnvData(env07, nc = 'srtm15plus', fileName = "D:/Environ_variables/depth07.nc", var = c('z'))
#env08 <- matchEnvData(env08, nc = 'srtm15plus', fileName = "D:/Environ_variables/depth08.nc", var = c('z'))
#env10 <- matchEnvData(env10, nc = 'srtm15plus', fileName = "D:/Environ_variables/depth10.nc", var = c('z'))
#env12 <- matchEnvData(env12, nc = 'srtm15plus', fileName = "D:/Environ_variables/depth12.nc", var = c('z'))
#env13 <- matchEnvData(env13, nc = 'srtm15plus', fileName = "D:/Environ_variables/depth13.nc", var = c('z'))
#env14 <- matchEnvData(env14, nc = 'srtm15plus', fileName = "D:/Environ_variables/depth14.nc", var = c('z'))
#env16 <- matchEnvData(env16, nc = 'srtm15plus', fileName = "D:/Environ_variables/depth16.nc", var = c('z'))
#env18 <- matchEnvData(env18, nc = 'srtm15plus', fileName = "D:/Environ_variables/depth18.nc", var = c('z'))
#env19 <- matchEnvData(env19, nc = 'srtm15plus', fileName = "D:/Environ_variables/depth19.nc", var = c('z'))
#env20 <- matchEnvData(env20, nc = 'srtm15plus', fileName = "D:/Environ_variables/depth20.nc", var = c('z'))
#env21 <- matchEnvData(env21, nc = 'srtm15plus', fileName = "D:/Environ_variables/depth21.nc", var = c('z'))
#env22 <- matchEnvData(env22, nc = 'srtm15plus', fileName = "D:/Environ_variables/depth22.nc", var = c('z'))
#env23 <- matchEnvData(env23, nc = 'srtm15plus', fileName = "D:/Environ_variables/depth23.nc", var = c('z'))
```

@@

Distance to Shore from ERDDAP using PAMPal function
```{r}
# Downloads data from servers onto external hard drive. Uncomment and only run once.
#env07 <- matchEnvData(env07, nc = 'dist2coast_1deg', fileName = "D:/Environ_variables/dist2shore07.nc", var = c('dist'))
#env08 <- matchEnvData(env08, nc = 'dist2coast_1deg', fileName = "D:/Environ_variables/dist2shore08.nc", var = c('dist'))
#env10 <- matchEnvData(env10, nc = 'dist2coast_1deg', fileName = "D:/Environ_variables/dist2shore10.nc", var = c('dist'))
#env12 <- matchEnvData(env12, nc = 'dist2coast_1deg', fileName = "D:/Environ_variables/dist2shore12.nc", var = c('dist'))
#env13 <- matchEnvData(env13, nc = 'dist2coast_1deg', fileName = "D:/Environ_variables/dist2shore13.nc", var = c('dist'))
#env14 <- matchEnvData(env14, nc = 'dist2coast_1deg', fileName = "D:/Environ_variables/dist2shore14.nc", var = c('dist'))
#env16 <- matchEnvData(env16, nc = 'dist2coast_1deg', fileName = "D:/Environ_variables/dist2shore16.nc", var = c('dist'))
#env18 <- matchEnvData(env18, nc = 'dist2coast_1deg', fileName = "D:/Environ_variables/dist2shore18.nc", var = c('dist'))
#env19 <- matchEnvData(env19, nc = 'dist2coast_1deg', fileName = "D:/Environ_variables/dist2shore19.nc", var = c('dist'))
#env20 <- matchEnvData(env20, nc = 'dist2coast_1deg', fileName = "D:/Environ_variables/dist2shore20.nc", var = c('dist'))
#env21 <- matchEnvData(env21, nc = 'dist2coast_1deg', fileName = "D:/Environ_variables/dist2shore21.nc", var = c('dist'))
#env22 <- matchEnvData(env22, nc = 'dist2coast_1deg', fileName = "D:/Environ_variables/dist2shore22.nc", var = c('dist'))
#env23 <- matchEnvData(env23, nc = 'dist2coast_1deg', fileName = "D:/Environ_variables/dist2shore23.nc", var = c('dist'))
```

@@

### Chlorophyll from ERDDAP using PAMPal function
Chlorophyll-a, Aqua MODIS, NPP, L3SMI, Global, 4km, Science Quality, 2003-
present (Monthly Composite) 
```{r}
# Downloads data from servers onto external hard drive. Uncomment and only run once.
#env07 <- matchEnvData(env07, nc = 'erdMH1chlamday', fileName = "D:/Environ_variables/chlorophyll07.nc", var = c('chlorophyll'))
#env08 <- matchEnvData(env08, nc = 'erdMH1chlamday', fileName = "D:/Environ_variables/chlorophyll08.nc", var = c('chlorophyll'))
#env10 <- matchEnvData(env10, nc = 'erdMH1chlamday', fileName = "D:/Environ_variables/chlorophyll10.nc", var = c('chlorophyll'))
#env12 <- matchEnvData(env12, nc = 'erdMH1chlamday', fileName = "D:/Environ_variables/chlorophyll12.nc", var = c('chlorophyll'))
#env13 <- matchEnvData(env13, nc = 'erdMH1chlamday', fileName = "D:/Environ_variables/chlorophyll13.nc", var = c('chlorophyll'))
#env14 <- matchEnvData(env14, nc = 'erdMH1chlamday', fileName = "D:/Environ_variables/chlorophyll14.nc", var = c('chlorophyll'))
#env16 <- matchEnvData(env16, nc = 'erdMH1chlamday', fileName = "D:/Environ_variables/chlorophyll16.nc", var = c('chlorophyll'))
#env18 <- matchEnvData(env18, nc = 'erdMH1chlamday', fileName = "D:/Environ_variables/chlorophyll18.nc", var = c('chlorophyll'))
#env19 <- matchEnvData(env19, nc = 'erdMH1chlamday', fileName = "D:/Environ_variables/chlorophyll19.nc", var = c('chlorophyll'))
#env20 <- matchEnvData(env20, nc = 'erdMH1chlamday', fileName = "D:/Environ_variables/chlorophyll20.nc", var = c('chlorophyll'))
#env21 <- matchEnvData(env21, nc = 'erdMH1chlamday', fileName = "D:/Environ_variables/chlorophyll21.nc", var = c('chlorophyll'))
#env22 <- matchEnvData(env22, nc = 'erdMH1chlamday', fileName = "D:/Environ_variables/chlorophyll22.nc", var = c('chlorophyll'))
#env23 <- matchEnvData(env23, nc = 'erdMH1chlamday', fileName = "D:/Environ_variables/chlorophyll23.nc", var = c('chlorophyll'))
```

@@

Sea Surface Height Anomalies from ERDDAP using PAMPal function
Sea Surface Height Anomalies from Altimetry, Global, 2017-
present (meters)
```{r}
# Downloads data from servers onto external hard drive. Uncomment and only run once.
#env07 <- matchEnvData(env07, nc = 'nesdisSSH1day', fileName = "D:/Environ_variables/ssh07.nc", var = c('sla'))
#env08 <- matchEnvData(env08, nc = 'nesdisSSH1day', fileName = "D:/Environ_variables/ssh08.nc", var = c('sla'))
#env10 <- matchEnvData(env10, nc = 'nesdisSSH1day', fileName = "D:/Environ_variables/ssh10.nc", var = c('sla'))
#env12 <- matchEnvData(env12, nc = 'nesdisSSH1day', fileName = "D:/Environ_variables/ssh12.nc", var = c('sla'))
#env13 <- matchEnvData(env13, nc = 'nesdisSSH1day', fileName = "D:/Environ_variables/ssh13.nc", var = c('sla'))
#env14 <- matchEnvData(env14, nc = 'nesdisSSH1day', fileName = "D:/Environ_variables/ssh14.nc", var = c('sla'))
#env16 <- matchEnvData(env16, nc = 'nesdisSSH1day', fileName = "D:/Environ_variables/ssh16.nc", var = c('sla'))
#env18 <- matchEnvData(env18, nc = 'nesdisSSH1day', fileName = "D:/Environ_variables/ssh18.nc", var = c('sla'))
#env19 <- matchEnvData(env19, nc = 'nesdisSSH1day', fileName = "D:/Environ_variables/ssh19.nc", var = c('sla'))
#env20 <- matchEnvData(env20, nc = 'nesdisSSH1day', fileName = "D:/Environ_variables/ssh20.nc", var = c('sla'))
#env21 <- matchEnvData(env21, nc = 'nesdisSSH1day', fileName = "D:/Environ_variables/ssh21.nc", var = c('sla'))
#env22 <- matchEnvData(env22, nc = 'nesdisSSH1day', fileName = "D:/Environ_variables/ssh22.nc", var = c('sla'))
#env23 <- matchEnvData(env23, nc = 'nesdisSSH1day', fileName = "D:/Environ_variables/ssh23.nc", var = c('sla'))
```

@@

#### HyCOM Data - MLDepth, MLTemp, TTemp, TDepth, Salinity

@@

```{r}
# Create loop for each buoy to download HYCOM data

## This only downloads 11 of 13 successfully - FIX
for (t in tracklist) {
  Hdata <- matchEnvData(as.data.frame(get(paste0("env", substr(t, 4, 5)))), nc=PAMmisc::hycomList, var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000)

@@

# runs hycom variable download and bind for one drift. can ignore if above works.
```{r}
## HYCOM function above stops at buoys 07 and 08, need to manually run it here. FIX!!
#test08 <- matchEnvData(as.data.frame(tracks_SS_08), nc=PAMmisc::hycomList, var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000)

#thermData08 <- bind_rows(lapply(test08, function(x) {
#    # do thermo calcs
#    result <- calcThermoVars(x$matchDepth, x$water_temp)
#    # get specific values
#    result$temp400 <- x$water_temp[x$matchDepth == 400]
#    result$sal400 <- x$salinity[x$matchDepth == 400]
#    result
#}))

#tracks_SS_08 <- cbind(tracks_SS_08, thermData08)
```

>>>>

Required Packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(sf)
library(terra)
library(PAMpal)
library(PAMmisc)
```

Read in GPSwhale rdas
```{r}
GPSwhaleAIS7 <- readRDS(file = 'D:/Buoy_dfs/GPSwhaleAIS07.rda')
GPSwhaleAIS8 <- readRDS(file = 'D:/Buoy_dfs/GPSwhaleAIS08.rda')
GPSwhaleAIS10 <- readRDS(file = 'D:/Buoy_dfs/GPSwhaleAIS10.rda')
GPSwhaleAIS12 <- readRDS(file = 'D:/Buoy_dfs/GPSwhaleAIS12.rda')
GPSwhaleAIS13 <- readRDS(file = 'D:/Buoy_dfs/GPSwhaleAIS13.rda')
GPSwhaleAIS14 <- readRDS(file = 'D:/Buoy_dfs/GPSwhaleAIS14.rda')
GPSwhaleAIS16 <- readRDS(file = 'D:/Buoy_dfs/GPSwhaleAIS16.rda')
GPSwhaleAIS18 <- readRDS(file = 'D:/Buoy_dfs/GPSwhaleAIS18.rda')
GPSwhaleAIS19 <- readRDS(file = 'D:/Buoy_dfs/GPSwhaleAIS19.rda')
GPSwhaleAIS20 <- readRDS(file = 'D:/Buoy_dfs/GPSwhaleAIS20.rda')
GPSwhaleAIS21 <- readRDS(file = 'D:/Buoy_dfs/GPSwhaleAIS21.rda')
GPSwhaleAIS22 <- readRDS(file = 'D:/Buoy_dfs/GPSwhaleAIS22.rda')
GPSwhaleAIS23 <- readRDS(file = 'D:/Buoy_dfs/GPSwhaleAIS23.rda')
```

@@

Sea Surface Temperature from ERDDAP using PAMPal function
Daily Multi-scale Ultra-high Resolution (MUR) SST Analysis fv04.1, Global, 0.01°
```{r}
# Downloads data from servers onto external hard drive. Uncomment and only run once.
env07 <- matchEnvData(GPSwhaleAIS7, nc = 'jplMURSST41', 
                      fileName = "D:/Environ_variables/analysedSST07", var = c('analysed_sst'))
env08 <- matchEnvData(GPSwhaleAIS8, nc = 'jplMURSST41', 
                      fileName = "D:/Environ_variables/analysedSST08", var = c('analysed_sst'))
env10 <- matchEnvData(GPSwhaleAIS10, nc = 'jplMURSST41', 
                      fileName = "D:/Environ_variables/analysedSST10", var = c('analysed_sst'))
env12 <- matchEnvData(GPSwhaleAIS12, nc = 'jplMURSST41', 
                      fileName = "D:/Environ_variables/analysedSST12", var = c('analysed_sst'))
env13 <- matchEnvData(GPSwhaleAIS13, nc = 'jplMURSST41', 
                      fileName = "D:/Environ_variables/analysedSST13", var = c('analysed_sst'))
env14 <- matchEnvData(GPSwhaleAIS14, nc = 'jplMURSST41', 
                      fileName = "D:/Environ_variables/analysedSST14", var = c('analysed_sst'))
env16 <- matchEnvData(GPSwhaleAIS16, nc = 'jplMURSST41', 
                      fileName = "D:/Environ_variables/analysedSST16", var = c('analysed_sst'))
env18 <- matchEnvData(GPSwhaleAIS18, nc = 'jplMURSST41', 
                      fileName = "D:/Environ_variables/analysedSST18", var = c('analysed_sst'))
env19 <- matchEnvData(GPSwhaleAIS19, nc = 'jplMURSST41', 
                      fileName = "D:/Environ_variables/analysedSST19", var = c('analysed_sst'))
env20 <- matchEnvData(GPSwhaleAIS20, nc = 'jplMURSST41', 
                      fileName = "D:/Environ_variables/analysedSST20", var = c('analysed_sst'))
env21 <- matchEnvData(GPSwhaleAIS21, nc = 'jplMURSST41', 
                      fileName = "D:/Environ_variables/analysedSST21", var = c('analysed_sst'))
env22 <- matchEnvData(GPSwhaleAIS22, nc = 'jplMURSST41', 
                      fileName = "D:/Environ_variables/analysedSST22", var = c('analysed_sst'))
env23 <- matchEnvData(GPSwhaleAIS23, nc = 'jplMURSST41', 
                      fileName = "D:/Environ_variables/analysedSST23", var = c('analysed_sst'))
```

@@

  assign(paste0("env", substr(ss, 33, 34)), data)
}
```

### Curl of Surface Wind Stress from ERDDAP using PAMPal function
FNMOC Ekman Transports, 1°, 6-hourly, Lon+/-180, Units: M Pa m-1
```{r}
# Downloads data from servers onto external hard drive. Uncomment and only run once.
#env07 <- matchEnvData(env07, nc = 'erdlasFnTran6_LonPM180', fileName = "D:/Environ_variables/windstress07.nc", var = c('curl'))
#env08 <- matchEnvData(env08, nc = 'erdlasFnTran6_LonPM180', fileName = "D:/Environ_variables/windstress08.nc", var = c('curl'))
#env10 <- matchEnvData(env10, nc = 'erdlasFnTran6_LonPM180', fileName = "D:/Environ_variables/windstress10.nc", var = c('curl'))
#env12 <- matchEnvData(env12, nc = 'erdlasFnTran6_LonPM180', fileName = "D:/Environ_variables/windstress12.nc", var = c('curl'))
#env13 <- matchEnvData(env13, nc = 'erdlasFnTran6_LonPM180', fileName = "D:/Environ_variables/windstress13.nc", var = c('curl'))
#env14 <- matchEnvData(env14, nc = 'erdlasFnTran6_LonPM180', fileName = "D:/Environ_variables/windstress14.nc", var = c('curl'))
#env16 <- matchEnvData(env16, nc = 'erdlasFnTran6_LonPM180', fileName = "D:/Environ_variables/windstress16.nc", var = c('curl'))
#env18 <- matchEnvData(env18, nc = 'erdlasFnTran6_LonPM180', fileName = "D:/Environ_variables/windstress18.nc", var = c('curl'))
#env19 <- matchEnvData(env19, nc = 'erdlasFnTran6_LonPM180', fileName = "D:/Environ_variables/windstress19.nc", var = c('curl'))
#env20 <- matchEnvData(env20, nc = 'erdlasFnTran6_LonPM180', fileName = "D:/Environ_variables/windstress20.nc", var = c('curl'))
#env21 <- matchEnvData(env21, nc = 'erdlasFnTran6_LonPM180', fileName = "D:/Environ_variables/windstress21.nc", var = c('curl'))
#env22 <- matchEnvData(env22, nc = 'erdlasFnTran6_LonPM180', fileName = "D:/Environ_variables/windstress22.nc", var = c('curl'))
#env23 <- matchEnvData(env23, nc = 'erdlasFnTran6_LonPM180', fileName = "D:/Environ_variables/windstress23.nc", var = c('curl'))

```
Once downloaded, matches wind stress curl from file
```{r}
windlist <- dir("D:/Environ_variables", recursive=TRUE, full.names=TRUE, pattern="windstress")

for (w in windlist) {
  data <- matchEnvData(get(paste0("env", substr(w, 32, 33))), nc = w) %>%
     dplyr::select(-matchLong_mean, -matchLat_mean, -matchTime_mean)   # remove unnecessary columns

@@

### Distance to continental slope (uses terra and sf package)
## Takes about 20 minutes to run
```{r}
bathy <- rast("D:/CCES_gebco_bathy.tif") # Bathymetry for all of CCE from https://download.gebco.net/
agg <- terra::aggregate(bathy, 5)  # Increase cell size by factor of 5 to be visible

@@

#Make sure to clear Global Environment of other objects with "env"
tracklist <- grep(pattern = "env", names(.GlobalEnv), value = T)

@@

  assign(paste0("env", substr(t, 4, 5)), dbind)
}
```
```{r} 
sf <- st_as_sf(get(t), coords = c("Longitude", "Latitude"), crs=4326, remove=F)
  trk <- terra::rasterize(vect(sf), agg)
  slop <- terra::distance(trk, vect(cont_slope))
  ext <- terra::extract(slop, vect(sf))

  dbind <- cbind(sf, ext) %>%
  dplyr::select(-ID) %>%
  rename("dist2slope" = last) %>%
  st_set_geometry(NULL) # remove geometry to prevent buoy dataframe from becoming sf object for future 
```

Depth from ERDDAP using PAMPal function
SRTM15_PLUS Estimated Topography, 15 seconds, Global, v1 
```{r}
# Downloads data from servers onto external hard drive. Uncomment and only run once.
#env07 <- matchEnvData(env07, nc = 'srtm15plus', fileName = "D:/Environ_variables/depth07.nc", var = c('z'))
#env08 <- matchEnvData(env08, nc = 'srtm15plus', fileName = "D:/Environ_variables/depth08.nc", var = c('z'))
#env10 <- matchEnvData(env10, nc = 'srtm15plus', fileName = "D:/Environ_variables/depth10.nc", var = c('z'))
#env12 <- matchEnvData(env12, nc = 'srtm15plus', fileName = "D:/Environ_variables/depth12.nc", var = c('z'))
#env13 <- matchEnvData(env13, nc = 'srtm15plus', fileName = "D:/Environ_variables/depth13.nc", var = c('z'))
#env14 <- matchEnvData(env14, nc = 'srtm15plus', fileName = "D:/Environ_variables/depth14.nc", var = c('z'))
#env16 <- matchEnvData(env16, nc = 'srtm15plus', fileName = "D:/Environ_variables/depth16.nc", var = c('z'))
#env18 <- matchEnvData(env18, nc = 'srtm15plus', fileName = "D:/Environ_variables/depth18.nc", var = c('z'))
#env19 <- matchEnvData(env19, nc = 'srtm15plus', fileName = "D:/Environ_variables/depth19.nc", var = c('z'))
#env20 <- matchEnvData(env20, nc = 'srtm15plus', fileName = "D:/Environ_variables/depth20.nc", var = c('z'))
#env21 <- matchEnvData(env21, nc = 'srtm15plus', fileName = "D:/Environ_variables/depth21.nc", var = c('z'))
#env22 <- matchEnvData(env22, nc = 'srtm15plus', fileName = "D:/Environ_variables/depth22.nc", var = c('z'))
#env23 <- matchEnvData(env23, nc = 'srtm15plus', fileName = "D:/Environ_variables/depth23.nc", var = c('z'))

```

@@

Distance to Shore from ERDDAP using PAMPal function
Distance to Nearest Coastline: 0.01-Degree Grid
```{r}
# Downloads data from servers onto external hard drive. Uncomment and only run once.
#env07 <- matchEnvData(env07, nc = 'dist2coast_1deg', fileName = "D:/Environ_variables/dist2shore07.nc", var = c('dist'))
#env08 <- matchEnvData(env08, nc = 'dist2coast_1deg', fileName = "D:/Environ_variables/dist2shore08.nc", var = c('dist'))
#env10 <- matchEnvData(env10, nc = 'dist2coast_1deg', fileName = "D:/Environ_variables/dist2shore10.nc", var = c('dist'))
#env12 <- matchEnvData(env12, nc = 'dist2coast_1deg', fileName = "D:/Environ_variables/dist2shore12.nc", var = c('dist'))
#env13 <- matchEnvData(env13, nc = 'dist2coast_1deg', fileName = "D:/Environ_variables/dist2shore13.nc", var = c('dist'))
#env14 <- matchEnvData(env14, nc = 'dist2coast_1deg', fileName = "D:/Environ_variables/dist2shore14.nc", var = c('dist'))
#env16 <- matchEnvData(env16, nc = 'dist2coast_1deg', fileName = "D:/Environ_variables/dist2shore16.nc", var = c('dist'))
#env18 <- matchEnvData(env18, nc = 'dist2coast_1deg', fileName = "D:/Environ_variables/dist2shore18.nc", var = c('dist'))
#env19 <- matchEnvData(env19, nc = 'dist2coast_1deg', fileName = "D:/Environ_variables/dist2shore19.nc", var = c('dist'))
#env20 <- matchEnvData(env20, nc = 'dist2coast_1deg', fileName = "D:/Environ_variables/dist2shore20.nc", var = c('dist'))
#env21 <- matchEnvData(env21, nc = 'dist2coast_1deg', fileName = "D:/Environ_variables/dist2shore21.nc", var = c('dist'))
#env22 <- matchEnvData(env22, nc = 'dist2coast_1deg', fileName = "D:/Environ_variables/dist2shore22.nc", var = c('dist'))
#env23 <- matchEnvData(env23, nc = 'dist2coast_1deg', fileName = "D:/Environ_variables/dist2shore23.nc", var = c('dist'))

```

@@

### Chlorophyll from ERDDAP using PAMPal function
Chlorophyll-a, Aqua MODIS, NPP, L3SMI, Global, 4km, Science Quality, 2003-
present (Monthly Composite)
Monthly is a better resolution because anything finer leaves NaN in columns
```{r}
# Downloads data from servers onto external hard drive. Uncomment and only run once.
#env07 <- matchEnvData(env07, nc = 'erdMH1chlamday', fileName = "D:/Environ_variables/chlorophyll07.nc", var = c('chlorophyll'))
#env08 <- matchEnvData(env08, nc = 'erdMH1chlamday', fileName = "D:/Environ_variables/chlorophyll08.nc", var = c('chlorophyll'))
#env10 <- matchEnvData(env10, nc = 'erdMH1chlamday', fileName = "D:/Environ_variables/chlorophyll10.nc", var = c('chlorophyll'))
#env12 <- matchEnvData(env12, nc = 'erdMH1chlamday', fileName = "D:/Environ_variables/chlorophyll12.nc", var = c('chlorophyll'))
#env13 <- matchEnvData(env13, nc = 'erdMH1chlamday', fileName = "D:/Environ_variables/chlorophyll13.nc", var = c('chlorophyll'))
#env14 <- matchEnvData(env14, nc = 'erdMH1chlamday', fileName = "D:/Environ_variables/chlorophyll14.nc", var = c('chlorophyll'))
#env16 <- matchEnvData(env16, nc = 'erdMH1chlamday', fileName = "D:/Environ_variables/chlorophyll16.nc", var = c('chlorophyll'))
#env18 <- matchEnvData(env18, nc = 'erdMH1chlamday', fileName = "D:/Environ_variables/chlorophyll18.nc", var = c('chlorophyll'))
#env19 <- matchEnvData(env19, nc = 'erdMH1chlamday', fileName = "D:/Environ_variables/chlorophyll19.nc", var = c('chlorophyll'))
#env20 <- matchEnvData(env20, nc = 'erdMH1chlamday', fileName = "D:/Environ_variables/chlorophyll20.nc", var = c('chlorophyll'))
#env21 <- matchEnvData(env21, nc = 'erdMH1chlamday', fileName = "D:/Environ_variables/chlorophyll21.nc", var = c('chlorophyll'))
#env22 <- matchEnvData(env22, nc = 'erdMH1chlamday', fileName = "D:/Environ_variables/chlorophyll22.nc", var = c('chlorophyll'))
#env23 <- matchEnvData(env23, nc = 'erdMH1chlamday', fileName = "D:/Environ_variables/chlorophyll23.nc", var = c('chlorophyll'))

```

@@

Sea Surface Height Anomalies from ERDDAP using PAMPal function
Sea Surface Height Anomalies from Altimetry, Global, 2017-
present (meters)
```{r}
# Downloads data from servers onto external hard drive. Uncomment and only run once.
#env07 <- matchEnvData(env07, nc = 'nesdisSSH1day', fileName = "D:/Environ_variables/ssh07.nc", var = c('sla'))
#env08 <- matchEnvData(env08, nc = 'nesdisSSH1day', fileName = "D:/Environ_variables/ssh08.nc", var = c('sla'))
#env10 <- matchEnvData(env10, nc = 'nesdisSSH1day', fileName = "D:/Environ_variables/ssh10.nc", var = c('sla'))
#env12 <- matchEnvData(env12, nc = 'nesdisSSH1day', fileName = "D:/Environ_variables/ssh12.nc", var = c('sla'))
#env13 <- matchEnvData(env13, nc = 'nesdisSSH1day', fileName = "D:/Environ_variables/ssh13.nc", var = c('sla'))
#env14 <- matchEnvData(env14, nc = 'nesdisSSH1day', fileName = "D:/Environ_variables/ssh14.nc", var = c('sla'))
#env16 <- matchEnvData(env16, nc = 'nesdisSSH1day', fileName = "D:/Environ_variables/ssh16.nc", var = c('sla'))
#env18 <- matchEnvData(env18, nc = 'nesdisSSH1day', fileName = "D:/Environ_variables/ssh18.nc", var = c('sla'))
#env19 <- matchEnvData(env19, nc = 'nesdisSSH1day', fileName = "D:/Environ_variables/ssh19.nc", var = c('sla'))
#env20 <- matchEnvData(env20, nc = 'nesdisSSH1day', fileName = "D:/Environ_variables/ssh20.nc", var = c('sla'))
#env21 <- matchEnvData(env21, nc = 'nesdisSSH1day', fileName = "D:/Environ_variables/ssh21.nc", var = c('sla'))
#env22 <- matchEnvData(env22, nc = 'nesdisSSH1day', fileName = "D:/Environ_variables/ssh22.nc", var = c('sla'))
#env23 <- matchEnvData(env23, nc = 'nesdisSSH1day', fileName = "D:/Environ_variables/ssh23.nc", var = c('sla'))

```

@@

#### Download HYCOM Data - MLDepth, MLTemp, TTemp, TDepth, Salinity

@@

```{r}
# Create loop for each buoy to download HYCOM data. Takes a long time to run.
for (t in tracklist) {
  Hdata <- matchEnvData(as.data.frame(get(paste0("env", substr(t, 4, 5)))), 
                        nc=PAMmisc::hycomList, 
                        var=c('water_temp', 'salinity'), 
                        raw=TRUE, 
                        depth = 0:1000)

@@

# runs hycom variable download and bind for one drift. Run if above does not work.
```{r}
Htest <- matchEnvData(env23, nc=PAMmisc::hycomList, var=c('water_temp', 'salinity'), raw=TRUE, depth = 0:1000)

thermData <- bind_rows(lapply(Htest, function(x) {
    # do thermo calcs
    result <- calcThermoVars(x$matchDepth, x$water_temp)
    # get specific values
    result$temp400 <- x$water_temp[x$matchDepth == 400]
    result$sal400 <- x$salinity[x$matchDepth == 400]
    result
}))

env23 <- cbind(env23, thermData)
```
